### 2020-02-26 overview.tex
To send for examination (oppo/comm).

Motivated by these, the present work embarks on a number of problems that originates from material fabrication, surface engineering and fluid transport.
We develop and utilise both high- and low-fidelity numerical models to simulate the underlying processes with the aims of
complementing experimental studies and elucidating their physical mechanisms.
The common ground at the bottom of these enquiries is that the processes all entail interactions of droplets in an ambient fluid or
are governed by the dynamics near a fluid interface.
Their specific backgrounds are given below (see also Chapter \ref{chp:misc}).

@@

\begin{figure}%[t]
  \centering
  \includegraphics[height=4.73cm]{photonics_cover.jpg}
  \includegraphics[height=4.73cm]{flow-assist1.png}
  \caption{(left) Cover photo of the January 2008 issue of Nature Photonics, showing an artistic rendering of a light beam passing through a 3D photonic crystal. \textcopyright \enspace Springer Nature. (right) Self-assembly of two to four droplets in a microfluidic chip; scale bar, 100 $\mu$m. Images courtesy of Dr.\ Bingqing Shen.}
  \label{fig:photonic-flow}
\end{figure}

@@

When particles are present in a fluid, the coupled dynamics is often significantly more complex than a single phase Stokes flow.
Theoretical analyses are thus limited to a handful of simple problems.
However, even in a simplified condition, an analytical solution often provides us useful physical intuitions that may be applied to more complicated cases.
Below, we list some elementary solutions of the Stokes equation in its simplest setup, \ie a single particle in unbounded Stokes flow.

@@

\begin{figure}%[t]
  \centering
  \includegraphics[width=0.49\columnwidth]{stokeslet1.pdf}
  \includegraphics[width=0.49\columnwidth]{stresslet1.pdf}
  \caption{Flow fields due a Stokeslet (left) and a stresslet (right) centered at the origin (red dot). The black arrows display the velocity magnitude, while their directions are also indicated by the gray arrows. The flow fields are both axi- and fore-aft symmetric. Note the difference in the velocity magnitude of the two flows.}
  \label{fig:lets}
\end{figure}

Finally, for a sphere of radius $a$, translating with velocity $\bm U$ and rotating with angular velocity $\bm \Omega$, its total drag, torque, and stresslet are given by the \emph{Fax\'{e}n's laws},
\begin{subequations} \label{eq:faxen}
 \begin{equation}
   {\bm F}^{d} = 6\pi\mu a\bigg( (1+\frac{a^2}{6}\nabla^2){\bm u^\infty - {\bm U}} \bigg) ,
 \end{equation}
 \begin{equation}
  {\bm T}^{d} = 8\pi\mu a^3 ({\bm \omega}^\infty - {\bm \Omega}),
 \end{equation}
 \begin{equation}
   {\bm S} = \frac{20}{3}\pi\mu a^3 \bigg( 1+\frac{a^2}{10}\nabla^2 \bigg) {\bm E}^\infty ,
 \end{equation}
\end{subequations}
where $\small {\bm E}^\infty=(\nabla {\bm u}^\infty + (\nabla {\bm u}^\infty)^T)/2$ is the rate-of-strain tensor.

@@

When particles are at close contact, say the surface gap $h \ll a$, with $a$ being the particle radius. Then, the hydrodynamic interaction will be dominated by the lubrication term. Specifically, the lubrication force diverge as $a/h$ in the longitudinal direction, and as $\log(a/h)$ in the transverse directions \citep{jeffrey_onishi_1984, jeffrey1992}. Qualitatively, lubrication functions as a dashpot that prevents incoming particles from colliding and separating particles from overshooting. These will be further discussed and demonstrated in Section \ref{sec:num-dem} and \emph{Paper 7}.

@@

Thirdly, we note that these observed particle interactions may be superimposed pairwise to infer the collective dynamics of a group of particles in similar conditions.
To illustrate this point, we simulated two rigid particles in a periodic channel (thus mimicking a 1D crystal) and
observed a crossover from permanent separation of the particle pair (for $0.32 \lessapprox z_0/H \lessapprox 0.68$)
to a trapping state (for $z_0/H \lessapprox 0.32$ or $z_0/H \gtrapprox 0.68$),
where $H$ is the channel height;%
\footnote{The computational setup and notations are the same as the case presented before, cf.\ Figures \ref{fig:itzhak-boris} and \ref{fig:stresslet-refl}.}
see Figure \ref{fig:xover}.
Qualitatively, the former results from cross-stream migrations that resemble the BG open trajectories,
while the latter is due to position swappings with neighbouring particles strongly influenced by the wall.
See \cite{Beatus2006, Janssen2012, Uspal2013} for a few more analogous and vivid demonstrations.

\begin{figure}%[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{H3.pdf}
  \caption{Evolution of a particle chain under various initial vertical positions (indicated by the color). The periodic channel has a height of $H=6$ and a length of $L=8$, both in units of particle radius. The solid and dashed lines denote the up- and downstream particles in the two-particle system, respectively. For $0.32 \lessapprox z_0/H \lessapprox 0.68$ (only a half plane is shown due to symmetry), the lines diverge (see the right panel) indicating a permanent separation.}
  \label{fig:xover}
\end{figure}

@@

\medskip
Consider the flow between two parallel plates where the gap $H$ is much smaller than the other two dimensions (say $x$ and $y$); such geometry is commonly known as the \emph{Hele Shaw cell}. Then, from the lubrication theory, the velocity of the flow within the thin layer is approximately prescribed as 
 \begin{equation} \label{eq:hele-shaw}
  \begin{aligned}
    {\bm u}(x,y,z) \approx -\frac{z(H-z)}{2\mu} \nabla p(x,y),
  \end{aligned}
 \end{equation}
where the vertical components of the velocity and pressure are neglected, though ${\bm u}$ remains $z$-dependent. Hence, the flow is quasi-two-dimensional.

@@

The Oseen tensor describing the long-range interactions within such a system is given as \citep{Leonardo_etal_2008, Diamant}
\begin{equation}
 \begin{aligned}
   {\bm B}^f({\bm x}) \approx \frac{1}{4\pi\mu H} \bigg[ \big( \ln(\nicefrac{1}{\kappa r}) -1 \big) {\bm I} +\frac{{\bm x}{\bm x} }{ \abs{{\bm x}}^2} \bigg],
   \quad \textrm{for} \quad \abs{\bm x} \gg H,
 \end{aligned}
\end{equation}
where $\kappa^{-1}$ is a cutoff length necessitated by the lateral system size or finite liquid inertia;
\ie the radial velocity component vanishes for $\abs{\bm x} \geq \kappa^{-1}$.%\footnote{By definition, $\kappa^{-1} \gg H$.}
Upon visualisation of the resulting flow field (Figure \ref{fig:q2d}, right),
it is obvious that particle interactions have a much longer range in free-standing liquid films.
Furthermore, the transverse motions remain strongly coupled at distances where the longitudinal mode has mostly decayed.
The slow logarithmic decay comparing to the dipolar flow in a Hele Shaw cell is physically resulting from the momentum conservation at the liquid boundaries,
as opposed to momentum absorption by the rigid plates \citep{Diamant}.
This difference may have implications to protein transport in biological membranes \citep{Saffman3111}.

@@

For a sedimenting suspension, the average velocity is given by
\begin{equation}
 \begin{aligned}
   \langle {\bm U} \rangle = \langle {\bm M} \rangle \cdot {\bm F},
 \end{aligned}
\end{equation}
where ${\bm M}$ is the \emph{mobility tensor} that relates the overall forcing ${\bm F}$, such as gravity, to the particle velocity ${\bm U}$, cf.\ Eq.\ \eqref{eq:stokeslet}; and $\langle \cdot \rangle$ denotes an average over all particles followed by an average over all configurations. Instantaneously, ${\bm M}$ depends only on the particle configuration instead of the velocity \citep{durlofsky_brady_bossis_1987}; its diagonal blocks represent the single-body drag coefficients, while its off-diagonal blocks embody the hydrodynamic interactions between the particles. In addition, for rigid particles undergoing purely hydrodynamic interactions, ${\bm M}$ is symmetric and positive definite due to the dissipative nature of the system \citep{graham_2018}. This property facilitates the proper inversion of ${\bm M}$ to produce the many-body approximation to the resistance matrix as required in some reduced-order computational models, see Section \ref{sec:num-dem}.

@@

One of the central tasks of rheology is to determine the bulk stress tensor $\langle \bm{\Sigma} \rangle_i$ of a flowing material, expressed as \citep{batchelor_1970, batchelor_1977}
\begin{equation} \label{eq:stress-tensor}
  \begin{aligned}
    \langle \bm{\Sigma} \rangle_i = -\bm{\Pi} +2\mu \bm{E}^\infty + \frac{N}{V} 
    \bigg( \big \langle \bm{S}^H  \big \rangle_i +  \big \langle \bm{S}^B  \big \rangle_i + \big \langle \bm{S}^A  \big \rangle_i \bigg), 
  \end{aligned}
\end{equation} 
where $V$ is the enclosing volume,
$N$ is the number of particles,
and $\bm{S}$'s denote the stresslets due to various particle interactions, including the hydrodynamic ($H$), Brownian ($B$), and any other applied ($A$) forces, as written.
Apart from the viscous stress tensor, $-\bm{\Pi} +2\mu \bm{E}^\infty$, due to the underlying flow, the rest of the terms in Eq.\ \eqref{eq:stress-tensor} are the particle contribution.
In general, these stress components are all related to the suspension microstructure.
In the case of the hydrodynamic stresslet, it can be computed as
\begin{equation} 
  \begin{aligned}
    \langle \bm{S}^H \rangle_i = - \langle \bm{R}_{SU}\cdot(\bm{U}-\bm{U}^\infty) - \bm{R}_{SE}:\bm{E}^\infty \rangle_i , 
  \end{aligned}
\end{equation}
where $\bm{R}_{SU}$ and $\bm{R}_{SE}$ are configuration-dependent resistance matrices that resemble $\bm{M}^{-1}$.
See Section \ref{subsec:sd} and \emph{Paper 7} for further details.

\bigskip
Suppose that the bulk stress tensor has been determined, then a number of \emph{effective} rheological properties can be readily extracted as follows.
Specifically, we focus on non-Brownian, rigid, neutrally buoyant spheres suspended in a Newtonian fluid with viscosity $\mu$ undergoing a simple shearing motion in the $xy$ plane with shear rate $\dot{\gamma}$.

@@

\medskip
When the suspension is non-dilute, normal stresses can deviate from each other as a result of the anisotropic microstructure or particle collision.
This leads to the definitions of the first and second normal stress differences as above, where $x$, $y$, $z$ denote the flow, velocity-gradient, and negative vorticity directions, respectively.
$N_1$ and $N_2$ are usually normalized by the modulus of the shear stress, $\abs{\bm{\Sigma}_{xy}}$, since they also diverge as $\phi \to \phi_J$.
The exact relationship between $N_1$ or $N_2$ and $\phi$ is unclear.
The overall picture is that $N_1$ is mostly of hydrodynamic origin, while $N_2$ is negative%
\footnote{Collision results in a negative normal stress because it is repulsive; the force is in the opposite direction to the position vector.}
and large in magnitude (\ie $|N_2| \gtrapprox |N_1|$) due to particle collisions, which happen more frequently in the $xy$ plane \citep{guazzelli_pouliquen_2018}.
See \cite{seto_giusteri_2018} for further discussions.

@@

The previous chapter depicts the physical pictures of droplet interactions and suspension flow in a few simplified conditions.
In the current chapter, we peek in the technical details of various numerical methods when theoretical analyses are insufficient and the full equations have to be simulated to describe the coupled fluid and particle dynamics.
First, we discuss standard Navier-Stokes solvers along with different interface-tracking/capturing methods. Simulations adopting these approaches intend to resolve all spatial and temporal scales within the system, and sometimes qualify as \emph{direct numerical simulations} (DNS).
Then, we discuss a special class of methods that solves the particle dynamics without directly computing the flow, thus is much faster, but becomes invalid beyond the Stokes flow approximation. Depending on the level of simplifications, these may fall under the category of \emph{coarse-grained} dynamics.
In view of the large amount of numerical methods and the constant development of computational fluid dynamics (CFD) in the last several decades, the list of methods is by no means complete. For more comprehensive surveys of available numerical techniques for multiphase flow problems, see \cite{prosperetti_tryggvason_2007, Rosti2019}.

@@

In order to evaluate the spatial derivatives in Eqs.\ \eqref{eq:projection-method}, the flow domain should be discretized so that either finite difference or weighted residual methods can be applied.
In essence, finite difference methods express the derivative as a truncated Taylor series expansion, while weighted residual methods assume that the solution can be represented analytically and they are usually formulated as one of the following better-known variants, including finite volume, finite element, or spectral methods \citep{Fletcher}.
Loosely speaking, finite difference and finite volume schemes are easier to implement on a structured grid; finite element methods are more flexible on non-Cartesian grids; and spectral methods have the highest order-of-accuracy.

Depending on the numerical condition and specific implementation, solving Eqs.\ \eqref{eq:projection-method} can be costly. The most computationally intensive part in a projection method is usually Eq.\ \eqref{eq:poisson}, where a Poisson equation for pressure must be solved numerically at each time step. Consequently, much work has gone into developing fast or iterative Poisson solvers that take advantage of fast Fourier transforms, Gauss elimination, or multigrid methods, see \cite{Buzbee_Golub_Nielson, Swarztrauber1977, multigrid_Brandt, Wesseling} for some early examples.

@@

Despite the conceptual simplicity, there are several technical details that require careful treatment when implementing front-tracking methods.
First, due to deformation, breakup or coalescence of the particle, or simply due to the non-uniform flow, the Lagrangian points may need to be redistributed, added, or deleted from time to time to ensure numerical stability and accuracy.
This is particularly cumbersome in a parallelised implementation.
Second, advection of the marker points requires interpolation of the velocity field, as the moving grid is generally non-aligned with the background fixed mesh.
Third, the internal boundary conditions (no-slip or surface tension) are to be constraint on the Eulerian mesh, thus a ``spreading'' procedure is necessary to feed the particle dynamics back to the flow. This is usually done with a smeared delta function.

@@

The feedback of the particle on the fluid flow are embedded in a diffuse body-forcing term, $\bm f$.
Depending on whether it is incorporated into the continuous equations before or after discretisation, IB methods can be broadly divided into continuous forcing and direct forcing.
Regardless of the approach, the spreading of the Lagrangian force onto the Eulerian mesh is realized with regularized Dirac delta functions, which may take different forms as long as $\sum_{ijk} \delta_p( \bm{x}_{ijk}-\bm{X}_{l}) \Delta x \Delta y \Delta z=1$ is satisfied for all $\bm{X}_{l}$.

@@

First proposed in the 1970s, volume-of-fluid (VOF) methods were among the oldest interface-capturing methods for modelling free-surface flow \citep{Debar1974, hirt_nichols_1975, hirt_nichols_1981a}, and they remain as being the more popular methods for multiphase flow simulations within the CFD community \citep{Scardovelli_Zaleski_1999, gerris, pilliod_puckett_2004, xiao_honma_kono_2005a, mthinc2012, paris2019}.
The principle idea is, first, introducing a cell-averaged volume fraction, $C_{k}$, defined as
\begin{equation}
 \begin{aligned}
   C_{k} = \frac{1}{\Delta V_k} \int_{\Omega_k} H(\bm{x}) dV,
 \end{aligned}
\end{equation}
where $\Delta V_k$ is the volume of computational cell $k$, and $H(\bm{x})$ is a phase indicator function, given as
\begin{equation}
    H(\bm{x}_k)=
    \begin{cases}
        1 \quad \quad & \textrm{if $\bm{x}$ is in the reference fluid,} \\
        0 \quad & \textrm{otherwise}, \\
    \end{cases}
\end{equation}
then, obtaining the temporal variation of the local volume fraction by solving the following equation,
\begin{equation} \label{eq:vof-advect}
 \begin{aligned}
   \frac{\partial C_{k}}{\partial t} + \int_{\partial \Omega_k} (\bm{u \cdot n}_k) H(\bm{x}) dS =0,
 \end{aligned}
\end{equation}
where $\bm{n}_k$ denote the six normal vectors (in a 3D Cartesian mesh) of surfaces $\partial \Omega_k$, and $\bm u$ is the local divergence-free velocity.

Note that, in the above formulations, the \emph{total} volume of each fluid phase is conserved by construction.
For many engineering calculations, this is the minimum requirement for a long-time simulation to be meaningful and makes VOF appealing in a wide range of industrial problems, such as predicting the impact of ocean waves splashing on the ship hull or designing spray nozzles to enhance fuel mixing in engines.
On the other hand, to correctly capture the local interface motion, the volume flux in Eq.\ \eqref{eq:vof-advect} must be calculated accurately, which requires certain information of the local interface \emph{geometry}, such as the normal direction and curvature.
This further requirement is not straightforward knowing only the cell-averaged volume fraction (see Figure \ref{fig:front-capture} for a very simplified illustration), therefore, many efforts have been devoted to developing methods that either reconstruct the interface in an extra step (these methods are called geometric VOF) or represent the interface in certain ways such that the volume flux can be computed directly (these methods are called algebraic VOF). See \cite{mirjalili_jain_dodd_2017a} for a recent overview of the different methods.

@@

If the particle inertia is negligible, the left-hand side of Eq.\ \eqref{eq:force-balance} goes to zero,
resulting in $\bm{F}^H \approx -\bm{F}^A$ and quasi-static particle dynamics.
If the fluid inertia is also negligible, the Navier-Stokes equations reduces to the Stokes equations and
$\bm{F}^H$ may be obtained without even explicitly solving the fluid motions.
In the following, we examine two such simplified techniques.
One is more accurate and holds for any particulate system in general, albeit computationally expensive. This is the so-called \emph{Stokesian dynamics}.
The other is derived from the former, but with additional features to model particle contact; hence, it is better suited for dense suspensions.
We dub it \emph{hybrid lubrication/granular dynamics}.

@@

From the fluid mechanics perspective, questions invoked by these findings naturally include,
(i) what mechanism leads to the drag reduction, and
(ii) how it can be optimised given all the constraints.
The constraints involve surface chemistry, contamination, degradation, fabrication defects, wetting, etc., making the problem truly multiscale and multiphysics.
Basic research along this line, including \emph{Paper 6}, is motivated to resolve at least part of these issues.

>>>>

Under such motivations, the present work embarks on a number of problems that originate from material fabrication, surface engineering and fluid transport.
We develop and utilise both high- and low-fidelity numerical models to simulate the underlying processes with the aims of
complementing experimental studies and elucidating their physical mechanisms.
The common ground at the bottom of these enquiries is that the processes all entail interactions of droplets in an ambient fluid or
are governed by the dynamics near a fluid interface.
Their specific backgrounds are given below (see also Chapter \ref{chp:misc}).

@@

\begin{figure}%[t]
  \centering
  \includegraphics[height=4.73cm]{photonics_cover.jpg}
  \includegraphics[height=4.73cm]{flow-assist1.png}
  \caption{(left) Cover photo of the January 2008 issue of Nature Photonics, showing an artistic rendering of a light beam passing through a 3D photonic crystal. \textcopyright \enspace Springer Nature. (right) Self-assembly of two to four droplets in a microfluidic chip; scale bars, 100 $\mu$m. Images courtesy of Dr.\ Bingqing Shen.}
  \label{fig:photonic-flow}
\end{figure}

@@

When particles are present in a fluid, the coupled dynamics is often significantly more complex than a single phase Stokes flow.
Theoretical analyses are thus limited to a handful of simple problems.
However, even in a simplified condition, analytical solutions often convey beautiful and strikingly useful physical intuitions that may guide us to understand the physics in more complicated cases.
Below, we list some elementary solutions of the Stokes equation in its simplest setup, \ie a single particle in unbounded Stokes flow.

@@

\begin{figure}%[t]
  \centering
  \includegraphics[width=0.49\columnwidth]{stokeslet1.pdf}
  \includegraphics[width=0.49\columnwidth]{stresslet1.pdf}
  \caption{Flow fields due a Stokeslet (left) and a stresslet (right) centered at the origin (red dot). The black arrows display the velocity magnitude, while their directions are also indicated by the gray arrows. The shapes of the flow field are both axi- and fore-and-aft symmetric. Note the difference in the velocity magnitude of the two flows.}
  \label{fig:lets}
\end{figure}

Finally, for a sphere of radius $a$, translating with velocity $\bm U$ and rotating with angular velocity $\bm \Omega$,
its total drag, torque and stresslet are given by the \emph{Fax\'{e}n's laws},
\begin{subequations} \label{eq:faxen}
 \begin{equation}
   {\bm F}^{d} = 6\pi\mu a\bigg( (1+\frac{a^2}{6}\nabla^2){\bm u^\infty - {\bm U}} \bigg) ,
 \end{equation}
 \begin{equation}
  {\bm T}^{d} = 8\pi\mu a^3 ({\bm \omega}^\infty - {\bm \Omega}),
 \end{equation}
 \begin{equation}
   {\bm S} = \frac{20}{3}\pi\mu a^3 \bigg( 1+\frac{a^2}{10}\nabla^2 \bigg) {\bm E}^\infty ,
 \end{equation}
\end{subequations}
where $\small {\bm E}^\infty=(\nabla {\bm u}^\infty + (\nabla {\bm u}^\infty)^T)/2$ is the rate-of-strain tensor.

@@

When particles are at close contact, say the surface gap $h \ll a$ with $a$ being the particle radius, the hydrodynamic interaction will be dominated by the lubrication term. Specifically, the lubrication force diverges as $a/h$ in the longitudinal direction, and as $\log(a/h)$ in the transverse directions \citep{jeffrey_onishi_1984, jeffrey1992}. Qualitatively, lubrication functions as a dashpot that prevents incoming particles from colliding and separating particles from overshooting. These will be further discussed and demonstrated in Section \ref{sec:num-dem} and \emph{Paper 7}.

@@

Thirdly, we note that these observed particle interactions may be superimposed pairwise to infer the collective dynamics of a group of particles in similar conditions.
To illustrate this point, we simulated two rigid particles in a periodic channel (thus mimicking a 1D crystal) and
observed a crossover from permanent separation of the particle pair (for $0.32 \lessapprox z_0/H \lessapprox 0.68$, \ie close to the mid-channel)
to a trapping state (for $z_0/H \lessapprox 0.32$ or $z_0/H \gtrapprox 0.68$, \ie close to the wall),
where $H$ is the channel height;%
\footnote{The computational setup and notations are the same as the case presented before, cf.\ Figures \ref{fig:itzhak-boris} and \ref{fig:stresslet-refl}.}
see Figure \ref{fig:xover}.
Qualitatively, the former results from cross-stream migrations that resemble the BG open trajectories,
while the latter is due to position swappings with neighbouring particles strongly influenced by the wall.
See \cite{Beatus2006, Janssen2012, Uspal2013} for a few more analogous and vivid demonstrations.

\begin{figure}%[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{H3.pdf}
  \caption{Evolution of a particle chain under various initial vertical positions (indicated by the color) in space-time coordinates. The periodic channel has a height of $H=6$ and a length of $L=8$, both in units of particle radius. The solid and dashed lines denote the up- and downstream particles in the two-particle system, respectively. For $0.32 \lessapprox z_0/H \lessapprox 0.68$ (only a half plane is shown due to symmetry), the lines diverge (see the right panel) indicating a permanent separation.}
  \label{fig:xover}
\end{figure}

@@

\medskip
Consider the flow between two parallel plates where the gap $H$ is much smaller than the other two dimensions (say $x$ and $y$); such geometry is commonly known as the \emph{Hele Shaw cell}. Then, from the lubrication theory, the velocity of the flow within the thin layer is approximately prescribed as 
 \begin{equation} \label{eq:hele-shaw}
  \begin{aligned}
    {\bm u}(x,y,z) \approx -\frac{z(H-z)}{2\mu} \nabla p(x,y),
  \end{aligned}
 \end{equation}
where the vertical components of the velocity and pressure are neglected, though ${\bm u}$ remains $z$-dependent. Hence, the flow is q2D.

@@

The Oseen tensor describing the long-range interactions within such a system is given as \citep{Leonardo_etal_2008, Diamant}
\begin{equation}
 \begin{aligned}
   {\bm B}^f({\bm x}) \approx \frac{1}{4\pi\mu H} \bigg[ \big( \ln(\nicefrac{1}{\kappa r}) -1 \big) {\bm I} +\frac{{\bm x}{\bm x} }{ \abs{{\bm x}}^2} \bigg],
   \quad \textrm{for} \quad \abs{\bm x} \gg H,
 \end{aligned}
\end{equation}
where $\kappa^{-1}$ is a cutoff length necessitated by the lateral system size or finite liquid inertia;
\ie the radial velocity component vanishes for $\abs{\bm x} \geq \kappa^{-1}$.%\footnote{By definition, $\kappa^{-1} \gg H$.}
Upon visualisation of the resulting flow field (Figure \ref{fig:q2d}, right),
it is obvious that particle interactions have a much longer range in free-standing liquid films.
Furthermore, the transverse motions remain strongly coupled at distances where the longitudinal mode has mostly decayed.
The slow logarithmic decay comparing to the dipolar flow in a Hele Shaw cell physically results from the momentum conservation at the liquid boundaries,
as opposed to momentum absorption by the rigid plates \citep{Diamant}.
This difference may have implications to protein transport in biological membranes \citep{Saffman3111}.

@@

For a sedimenting suspension, the average velocity is given by
\begin{equation}
 \begin{aligned}
   \langle {\bm U} \rangle = \langle {\bm M} \cdot {\bm F} \rangle,
 \end{aligned}
\end{equation}
where ${\bm M}$ is the \emph{mobility tensor} that relates the overall forcing ${\bm F}$, such as gravity, to the particle velocity ${\bm U}$, cf.\ Eq.\ \eqref{eq:stokes-green}; and $\langle \cdot \rangle$ denotes an average over all particles followed by an average over all configurations. Instantaneously, ${\bm M}$ depends only on the particle configuration instead of the velocity \citep{durlofsky_brady_bossis_1987}; its diagonal blocks represent the single-body drag coefficients, while its off-diagonal blocks embody the hydrodynamic interactions between the particles. In addition, for rigid particles undergoing purely hydrodynamic interactions, ${\bm M}$ is symmetric and positive definite due to the dissipative nature of the system \citep{graham_2018}. This property facilitates the proper inversion of ${\bm M}$ to produce the many-body approximation to the resistance matrix as required in some reduced-order computational models, see Section \ref{sec:num-dem}.

@@

One of the central tasks of rheology is to determine the bulk stress tensor $\bm{\Sigma}$ of a flowing material, expressed as \citep{batchelor_1970, batchelor_1977}
\begin{equation} \label{eq:stress-tensor}
  \begin{aligned}
    \bm{\Sigma} = -\bm{\Pi} +2\mu \bm{E}^\infty + \frac{N}{V} 
    \bigg( \big \langle \bm{S}^H  \big \rangle_i +  \big \langle \bm{S}^B  \big \rangle_i + \big \langle \bm{S}^A  \big \rangle_i \bigg), 
  \end{aligned}
\end{equation} 
where $V$ is the enclosing volume,
$N$ is the number of particles,
and $\bm{S}$'s denote the stresslets due to various particle interactions, including the hydrodynamic ($H$), Brownian ($B$), and any other applied ($A$) forces, as written.
Apart from the viscous stress tensor, $-\bm{\Pi} +2\mu \bm{E}^\infty$, due to the underlying flow, the rest of the terms in Eq.\ \eqref{eq:stress-tensor} are the particle contribution.
In general, these stress components are all related to the suspension microstructure.
In the case of the hydrodynamic stresslet, it can be computed as
\begin{equation} 
  \begin{aligned}
    \langle \bm{S}^H \rangle_i = - \langle \bm{R}_{SU}\cdot(\bm{U}-\bm{U}^\infty) - \bm{R}_{SE}:\bm{E}^\infty \rangle_i , 
  \end{aligned}
\end{equation}
where $\bm{R}_{SU}$ and $\bm{R}_{SE}$ are configuration-dependent resistance matrices that resemble $\bm{M}^{-1}$.
See Section \ref{subsec:sd} and \emph{Paper 7} for further details.

\medskip
Suppose that the bulk stress tensor has been determined, then a number of effective properties characterising the suspension rheology can be readily extracted.
In the following, we define three such quantities in a specific case -- non-Brownian, rigid and neutrally buoyant spheres suspended in a Newtonian fluid with viscosity $\mu$ undergoing a simple shearing motion in the $xy$ plane with shear rate $\dot{\gamma}$.

@@

\medskip
When the suspension is non-dilute, normal stresses can deviate from each other as a result of the anisotropic microstructure or particle collision.
This leads to the definitions of the first and second normal stress differences as above, where $x$, $y$, $z$ denote the flow, velocity-gradient, and negative vorticity directions, respectively.
$N_1$ and $N_2$ are usually normalized by the modulus of the shear stress, $\abs{\bm{\Sigma}_{xy}}$, since they also diverge as $\phi \to \phi_J$.
The exact relationship between $N_1$ or $N_2$ and $\phi$ is unclear.
The overall picture is that $N_1$ is mostly of hydrodynamic origin, while $N_2$ is negative%
\footnote{Collision results in a negative normal stress because it is repulsive; the force is in the opposite direction to the position vector.}
and large in magnitude (\ie $|N_2| \gtrapprox |N_1|$) due to particle collisions, which happen more frequently in the $xy$ plane in the dense regime \citep{guazzelli_pouliquen_2018}.
See \cite{seto_giusteri_2018} for further discussions.

@@

The previous chapter depicts the physical pictures of droplet interactions and suspension flow in a few simplified conditions.
In the current chapter, we peek in the technical details of various numerical methods when theoretical analyses are insufficient and the full equations have to be simulated to describe the coupled fluid and particle dynamics.
First, we discuss standard Navier-Stokes solvers along with different interface-tracking/capturing methods. Simulations adopting these approaches intend to resolve all spatial and temporal scales within the system, and sometimes qualify as \emph{direct numerical simulations} (DNS).
Then, we discuss a special class of methods that solves the particle dynamics without directly computing the flow, thus is much faster but only valid in the Stokes flow regime. Depending on the level of simplifications, these may fall under the category of \emph{coarse-grained} dynamics.
In view of the large amount of numerical methods and the constant development of computational fluid dynamics (CFD) in the last several decades, the list of methods is by no means complete. For more comprehensive surveys of available numerical techniques for multiphase flow problems, see \cite{prosperetti_tryggvason_2007, Rosti2019}.

@@

In order to evaluate the spatial derivatives in Eqs.\ \eqref{eq:projection-method},
the flow domain should be discretized so that either finite difference or weighted residual methods can be applied.
In essence, finite difference methods express the derivative as a truncated Taylor series expansion,
while weighted residual methods assume that the solution can be represented analytically and they are usually formulated as one of the following better-known variants
including finite volume, finite element or spectral methods \citep{Fletcher}.
Loosely speaking, finite difference and finite volume schemes are easier to implement on a structured grid;
finite element methods are more flexible on non-Cartesian grids; and spectral methods have the highest order-of-accuracy.

Depending on the numerical condition and specific implementation, solving Eqs.\ \eqref{eq:projection-method} can be costly.
The most computationally intensive part in a projection method is usually Eq.\ \eqref{eq:poisson}, where a Poisson equation for pressure must be solved numerically at each time step.
Consequently, much work has gone into developing fast or iterative Poisson solvers that take advantage of fast Fourier transforms, Gauss elimination or multigrid methods,
see \cite{Buzbee_Golub_Nielson, Swarztrauber1977, multigrid_Brandt, Wesseling} for some early examples.

@@

Despite the conceptual simplicity, there are several technical details that require careful treatment when implementing front-tracking methods.
First, due to deformation, breakup or coalescence of the particle, or simply due to the non-uniform flow, the Lagrangian points may need to be redistributed, added or deleted from time to time to ensure numerical stability and accuracy.
This is particularly cumbersome in a parallelised implementation.
Second, advection of the marker points requires interpolation of the velocity field as the moving grid is generally non-aligned with the background fixed mesh.
Third, the internal boundary conditions (no-slip or surface tension) are to be imposed on the Eulerian mesh, thus a ``spreading'' procedure is necessary to feed the particle dynamics back to the flow. This is usually done with a smeared delta function.

@@

The feedback of the particle on the fluid flow are embedded in a diffuse body-forcing term, $\bm f$.
Depending on whether it is incorporated into the continuous equations before or after discretization, IB methods can be broadly divided into continuous forcing and direct forcing.
Regardless of the approach, the spreading of the Lagrangian force onto the Eulerian mesh is realized with regularized Dirac delta functions, which may take different forms as long as $\sum_{ijk} \delta_p( \bm{x}_{ijk}-\bm{X}_{l}) \Delta x \Delta y \Delta z=1$ is satisfied for all $\bm{X}_{l}$.

@@

First proposed in the 1970s, volume-of-fluid (VOF) methods were among the oldest interface-capturing methods for modelling free-surface flow \citep{Debar1974, hirt_nichols_1975, hirt_nichols_1981a}, and they remain as being the more popular methods for multiphase flow simulations within the CFD community \citep{Scardovelli_Zaleski_1999, gerris, pilliod_puckett_2004, xiao_honma_kono_2005a, mthinc2012, paris2019}.
The principle idea is to first introduce a cell-averaged volume fraction, $C_{k}$, defined as
\begin{equation}
 \begin{aligned}
   C_{k} = \frac{1}{\Delta V_k} \int_{\Omega_k} H(\bm{x}) dV,
 \end{aligned}
\end{equation}
where $\Delta V_k$ is the volume of computational cell $k$, and $H(\bm{x})$ is a phase indicator function, given as
\begin{equation}
    H(\bm{x}_k)=
    \begin{cases}
        1 \quad \quad & \textrm{if $\bm{x}$ is in the reference fluid,} \\
        0 \quad & \textrm{otherwise}, \\
    \end{cases}
\end{equation}
then obtain the temporal variation of the local volume fraction by solving the following equation,
\begin{equation} \label{eq:vof-advect}
 \begin{aligned}
   \frac{\partial C_{k}}{\partial t} + \int_{\partial \Omega_k} (\bm{u \cdot n}_k) H(\bm{x}) dS =0,
 \end{aligned}
\end{equation}
where $\bm{n}_k$ denote the six normal vectors (in a 3D Cartesian mesh) of surfaces $\partial \Omega_k$, and $\bm u$ is the local divergence-free velocity.

Note that, in the above formulations, the total volume of each fluid phase is conserved by construction.
For many engineering calculations, this is the minimum requirement for a long-time simulation to be meaningful and makes VOF appealing in a wide range of industrial problems, such as predicting the impact of ocean waves splashing on the ship hull or designing spray nozzles to enhance fuel mixing in engines.
On the other hand, to correctly capture the local interface motion, the volume flux in Eq.\ \eqref{eq:vof-advect} must be calculated accurately, which requires certain information of the local interface geometry, such as the normal direction and curvature.
This further requirement is not straightforward knowing only the cell-averaged volume fraction (see Figure \ref{fig:front-capture} for a very simplified illustration), therefore, many efforts have been devoted to developing methods that either reconstruct the interface in an extra step (these methods are called geometric VOF) or represent the interface in certain ways such that the volume flux can be computed directly (these methods are called algebraic VOF). See \cite{mirjalili_jain_dodd_2017a} for a recent overview of the different methods.

@@

If the particle inertia is negligible, the left-hand side of Eq.\ \eqref{eq:force-balance} goes to zero,
resulting in $\bm{F}^H \approx -\bm{F}^A$ and quasi-static particle dynamics.
If the fluid inertia is also negligible, the Navier-Stokes equations reduce to the Stokes equations and
$\bm{F}^H$ may be obtained without even explicitly solving the fluid motions.
In the following, we examine two such simplified techniques.
One is more accurate and holds for any particulate system in general, albeit computationally expensive. This is the so-called \emph{Stokesian dynamics}.
The other is derived from the former, but with additional features to model particle contact; hence, it is better suited for dense suspensions.
We dub it \emph{hybrid lubrication/granular dynamics}.

@@

From the fluid mechanics perspective, questions invoked by these findings naturally include:
(i) what mechanism leads to the drag reduction, and
(ii) how it can be optimised given all the constraints.
The constraints involve surface chemistry, contamination, degradation, fabrication defects, wetting, etc., making the problem truly multiscale and multiphysics.
Basic research along this line, including \emph{Paper 6}, is motivated to resolve at least part of these issues.
